---
title: "Data Reformatting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(tidyr)

# to call stats filter - stats::filter()
```


# Reformat Catch Data

* remove "all" column
* create species column
    - move from wide to long
* general QA

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(catch_original)
```

Remove the `all` and `notesRegCode` columns using `select`
Cmd + shift + m : pipe operator shortcut %>% 

```{r}
catch_long <- catch_original %>% 
    select(-All, -notesRegCode) %>% # select columns we want
    gather(key = species, value = catch, #move from wide to long
           -Region, -Year)

head(catch_long)
```

```{r}
catch_wide <- catch_long %>% 
    spread(key = Region, value = catch)

head(catch_wide)
```

## Clean up our data

* rename catch to catch_thousands
* change catch column to numeric
* create a new catch column in units num. of fish

```{r}
catch_clean <- catch_long %>% 
    rename(catch_thousands = catch) %>% 
    mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
    mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
    mutate(catch = catch_thousands * 1000) %>% 
    select(-catch_thousands)

head(catch_clean)
```


```{r, eval = F, echo = F}
#Look for where as.numeric failed
test <- as.numeric(catch_long$catch)
i <- which(is.na(test) == TRUE)
catch_long[i,]
```

## Split - Apply - Combine

* calculate mean catch by species

```{r}
species_mean <- catch_clean %>% 
    group_by(species, Region) %>% 
    summarise(catch_mean = mean(catch),
              num_obs = length(which(is.na(catch) == F)))

head(species_mean)
```

```{r}
year_mean <- catch_clean %>% 
    group_by(Year) %>% 
    summarize(catch_mean = mean(catch)) %>% 
    arrange(-Year) %>% 
    filter(Year <= 1990 & Year >= 1960) # |

year_mean
```



# Join the Region Definitions

Read in the regions dataframe

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea) %>% 
    rename(Region = code)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_clean, region_defs, by = c("Region"))

head(catch_joined)
```


# Misc. Functions

Separate and unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)
```

```{r}
dates_split <- dates_df %>% 
    separate(date, into = c("month", "day", "year"), sep = "/", remove = F)

head(dates_split)
```


```{r}
dates_unite <- dates_split %>% 
    unite(col = date_iso, year, month, day, sep = "-")

head(dates_unite)
```

```{r}
isls <- data.frame(site = c("ALA-02",
                                "ALA-03-12",
                                "GUG-12",
                                "GUA-21",
                                "GUAg-12"),
                       stringsAsFactors = FALSE)

isls <- isls %>% 
    separate(site, into = c("island", "site"))

isls
```

